AHABus \& FCORE: A Modular Bus System for High-Altitude Balloon Missions

### 2017-04-18 chapters/03-methodology.tex
Add radio discussion
Add bus discussion
First draft, minus photos

@@

The proposed solution for this project is inspired by the bus model that
supports the architecture of most commercial satellites. A standard software
and hardware platform provides common facilities required on all High-Altitude
flights – position tracking, internal data transfers and radio telemetry to the
ground station – allowing payloads (scientific instruments for example) to
only generate data and make it available to the platform.

@@

The proposed solution was designed to run on low-cost embedded platforms, and
the prototype was built around the Espressif ESP8266 micro-controller
(\acrlong{obc}). The radio link used a Radiometrix NTX2B transmitter
(\cite{radiometrix2012}), and an unnamed Software-Defined Radio receiver and
antenna.

@@

RTTY itself functions in the same way as wired UART serial links. It can provide
some level of error checking if the transmission of a parity bit is enabled, but
it does not carry enough information to correct errors. Additionally, RTTY is
purely a link layer, and does not provide any kind of encapsulation necessary
for multiplexing data from different payloads.

\subsection{Hardware Setup}

@@

\begin{enumerate}
\item When a payload is ready to send data, it pulls its $\overline{CR}$ line
low.
\item The \acrshort{obc} lists payloads that have raised CR. For each:
\begin{enumerate}
    \item The \acrshort{obc} pulls the payload's $\overline{CE}$ line low.
    \item The payload sends its data to the \acrshort{obc}.
    \item The \acrshort{obc} release the payload's $\overline{CE}$ line to high.
    \item The payload releases its $\overline{CR}$ line to high.
    \item The \acrshort{obc} processes the received data.
\end{enumerate}
\end{enumerate}

@@

A Bus Tasks runs at a fixed interval, configured at compile time depending on
the payload's priority level, as discussed in more details in section
\ref{sssec:conf-file}. The task establishes a connection with the payload using
the \acrshort{ahabus} payload bus protocol. If data is available, it is
collected into a temporary buffer and a packet with the platform's cached
geographical coordinates is created and sent to the radio submodule.

@@

\[B = F_S + n_{1}F_{P} + \frac{n_{2}F_{P}}{2} + \frac{n_{3}F_{P}}{3} \]

>>>>

\section{Introduction}

In this chapter, the \acrfull{ahabus} is presented as a prototype solution to the issues exposed in chapter \ref{ch:lit-review}.

@@

The proposed solution for this project is inspired by the bus model that
supports the architecture of most commercial satellites. A standard software
and hardware platform provides common facilities required on all High-Altitude
flights – position tracking, internal data transfers and radio telemetry to the
ground station – allowing payloads (scientific instruments for example) to
only generate data and make it available to the platform. A global overview of the platform and the environment used for its development is presented, followed by a detailed discussion of the development process of each of the four main components of \acrshort{ahabus} (Radio Communications, Payload Bus, Flight Software and Ground Segment). Finally, the methods used to evaluate the viability of the prototype platform are presented, leading to the results presented in chapter \ref{ch:results}.

The source code and protocol documentation resulting from the development process discussed in this chapter are available in the project's repositories listed in appendix \ref{apx:source-repos}.

@@

The proposed solution was designed to run on low-cost embedded platforms, and the prototype was built around the Espressif ESP8266 (\cite{EspressifInc.2017}) micro-controller (\acrlong{obc}). The radio link used a Radiometrix NTX2B transmitter (\cite{radiometrix2012}), and an unnamed Software-Defined Radio receiver and antenna.

@@

RTTY itself functions in the same way as wired UART serial links. It can provide
some level of error checking if the transmission of a parity bit is enabled, but
it does not carry enough information to correct errors. Additionally, RTTY is
purely a link layer, and does not provide any kind of encapsulation necessary
for multiplexing data from different payloads.
%
% \subsection{Hardware Setup}
%
% The radio tra

@@

\begin{enumerate}
\item When a payload is ready to send data, it pulls its $\overline{CR}$ line
low.
\item The \acrshort{obc} lists payloads that have raised CR. For each:
\begin{enumerate}
\item The \acrshort{obc} pulls the payload's $\overline{CE}$ line low.
\item The payload sends its data to the \acrshort{obc}.
\item The \acrshort{obc} releases the payload's $\overline{CE}$ line to high.
\item The payload releases its $\overline{CR}$ line to high.
\item The \acrshort{obc} processes the received data.
\end{enumerate}
\end{enumerate}

@@

A Bus Tasks runs at a fixed interval, configured at compile time depending on
the payload's priority level, as discussed in more details in section
\ref{ssec:conf-file}. The task establishes a connection with the payload using
the \acrshort{ahabus} payload bus protocol. If data is available, it is
collected into a temporary buffer and a packet with the platform's cached
geographical coordinates is created and sent to the radio submodule.

@@

\[B = F_S + n_{1}F_{P} + n_{2}\frac{F_{P}}{2} + n_{3}\frac{F_{P}}{3} \]

### 2017-04-18 chapters/04-results.tex
Add stress tests discussion
First draft, minus photos

The following section presents the observations and numerical results obtained
during the tests done as part of the evaluation process of the project, as
discussed in section \ref{sec:testing}.

@@

\end{tabular}
\end{center}
\caption {DITL Test - Received packets}
\label{tab:ditl-health}
\end{table}

@@

When the average byte error rate, as shown in figure \ref{fig:stress-14.256},
some frames can be corrected, while some where the error rate is higher than
the average are lost. The fluctuations in actual error rate are due to the use
of the C language's \texttt{rand()} function. %remote or explain

>>>>

The following section presents the observations and numerical results obtained
during the tests done as part of the evaluation process of the project, as
discussed in section \ref{sec:testing}. The raw data files from which the
results presented in this chapter are derived, as well as the Python scripts
used to process them, are available in the project's repositories listed in
appendix \ref{apx:test-data-repos}.

@@

\end{tabular}
\end{center}
\caption {DITL Test - Received packets}
\label{tab:ditl-packet-count}
\end{table}

@@

When the average byte error rate, as shown in figure \ref{fig:stress-14.256},
some frames can be corrected, while some where the error rate is higher than
the average are lost.

### 2017-04-18 chapters/05-discussion.tex
First draft, minus photos

The aim of the project was to design and build a prototype of a High-Altitude
Balloon modular platform, similar in principle to buses used in the satellite
industry, and evaluate both the feasibility of such a platform and its
performance. The platform was required to allow for stable internal
communications between the payloads and the \acrlong{obc}, tracking of
ancillary data (\acrshort{gnss} co-ordinates and altitude) and reliable
transmission of the collected telemetry to one or multiple ground stations 
through a low-power radio channel. In this chapter, the \acrshort{ahabus}
platform prototype is evaluated against these aims, first by analysing the
results of tests discussed in chapter \ref{ch:results}, then through more
general qualitative observation and analysis.%  This chapter finally presents the
% further work that would be required for the platform to fully meet its design
% goals, leading to potential test flights.

@@

This section aims to explain and analyse the quantitative results obtained
during the testing phase of the project. The results are presented in chapter
\ref{ch:results}, and the raw data files obtained from each tests are included
in appendices \ref{} through \ref{}, or on the project's GitHub account at
https://github.com/ahabus/results

>>>>

The aim of the project was to design and build a prototype of a High-Altitude
Balloon modular platform, similar in principle to buses used in the satellite
industry, and evaluate both the feasibility of such a platform and its
performance. The platform was required to allow for stable internal
communications between the payloads and the \acrlong{obc}, tracking of
ancillary data (\acrshort{gnss} co-ordinates and altitude) and reliable
transmission of the collected telemetry to one or multiple ground stations 
through a low-power radio channel. In this chapter, the \acrshort{ahabus}
platform prototype is evaluated against these aims, firs through general 
qualitative observations and analysis of the prototype's behaviour, then by
analysing the results of tests discussed in chapter \ref{ch:results}.

@@

This section aims to explain and analyse the quantitative results obtained
during the testing phase of the project. The results are presented in chapter
\ref{ch:results}, and the raw data files obtained from each tests are made available in the repositories specified in appendix \ref{apx:test-data-repos}.

### 2017-04-18 paper.tex
First draft, minus photos

\documentclass[11pt]{report}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{titlesec}
\usepackage{float}
\usepackage{color}
\usepackage{listings}
\usepackage{lmodern}
\usepackage{graphicx}
\usepackage{bytefield}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{csvsimple}
\usepackage{hyperref}

@@

\usepackage[detect-weight=true, binary-units=true]{siunitx}
\usepackage{tikz}
\usepackage[justification=centering,labelfont={small,bf},font={small,bf}]{caption}
\usepackage[scaled]{beramono} % BitStream Vera font (code)
\usepackage[T1]{fontenc}
\usepackage[a4paper, margin=3cm]{geometry}
\usepackage[nottoc]{tocbibind}
\usepackage{array}

@@

\begin{abstract}
\end{abstract}

@@

>>>>

\documentclass[11pt]{report}
\usepackage[scaled]{beramono} % BitStream Vera font (code)
\usepackage[T1]{fontenc}
\usepackage[a4paper, margin=3cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{titlesec}
\usepackage{float}
\usepackage{color}
\usepackage{listings}
\usepackage{lmodern}
\usepackage{graphicx}
\usepackage{bytefield}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{csvsimple}
\usepackage{hyperref}

@@

\usepackage[detect-weight=true, binary-units=true]{siunitx}
\usepackage{tikz}
\usepackage[justification=centering,labelfont={small,bf},font={small,bf}]{caption}
\usepackage[titletoc,toc,page]{appendix}
\usepackage[nottoc]{tocbibind}
\usepackage{array}

@@

\begin{abstract}

The aim of this project was to investigate the design options for a modular payload bus for \acrfull{hab} missions, and to evaluate one such design by implementing in in hardware and software.

\acrlong{hab} provide affordable access to near-space conditions, even for low-budget entities (universities, schools and amateurs). While current low-budget missions rely on commercial off-the-shelf computers and sensors, their architecture and software are often custom-made and targeted for a single use. This prevents easy reuse of hardware and software components, and raises the barrier to entry by requiring non-trivial network and computing skills.

This paper presents the development of \acrfull{ahabus}, solution designed by applying the modular bus model used by most modern satellites to \acrlong{hab} architectures. A common, generic \acrfull{obc} running a common flight software (\acrshort{fcore}) provides facilities for position tracking and reliable radio telemetry to a ground station, and polls data from payloads, custom for each flight.

The resulting protocols and platform prototype were successful and passed both day-in-the-life testing and radio channel resilience tests. Further testing of the and minor modifications to the Payload Bus and Radio Communications protocols will be required in order to qualify the platform for in-flight use. Once these steps are taken, \acrshort{ahabus} could render scientific \acrlong{hab} missions more accessible and affordable by allowing students and researchers to focus on payload design and development.

\end{abstract}

@@

\input{appendices.tex}
